---
title: "Understand design from Ostarek & Huetting (2017) and Lupyan & Ward (2013)"
author: '[Guillermo Montero-Melis](https://www.mpi.nl/people/montero-melis-guillermo)'
date: '`r as.character(format(Sys.Date(), format="%d/%m/%Y"))`'
output:
  html_document:
    depth: 2
    number_sections: yes
    theme: default
    toc: yes
---


Setup workspace
==============

Load libraries

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE}
library("knitr")
library("tidyverse")  # ggplot2, dplyr, readr, purrr, etc
theme_set(theme_bw())
```


Load data sets:

```{r, message=FALSE}
# Load data OH17
ost <- read_csv("data_ostarek2017.csv")
head(ost) %>% kable
# LW13
lw1 <- read_csv("data_lupyan13_exp1_simple.csv")
head(lw1) %>% kable
```

OH17
====

The design in OH17 is very clean and balanced, and straightforward to understand.

Number of participants:

```{r}
length(unique(ost$participant))
```


Unique target images and their frequency.

```{r}
table(ost$target)
```

Unique primes and their frequency.

```{r}
table(ost$prime)
```


Prime-target combinations

```{r}
ost %>%
  group_by(target, prime, condition) %>%
  summarise(N = n()) %>%
  arrange(target, condition, prime) %>%
  kable
```


There's a slight imbalance for some primes, but it is present across the board,
not just for a few participants:

```{r}
with(ost, addmargins(table(participant, prime)), 1)
```

Which are the odd items in the design?

```{r}
ost %>%
  group_by(participant, target, prime, condition) %>%
  summarise(N = n()) %>%
  filter(! N %in% c(1, 4, 8)) %>%
  arrange(participant, target, condition, prime) %>%
  kable
```



LW13 (Exp 1)
====

Basics
------

The data frame:

```{r}
lw1 %>% head %>% kable
```


Number of participants:

```{r}
length(unique(lw1$participant))
```


Unique target images and their frequency.

```{r}
sort(table(lw1$target, useNA = "ifany"))
```

Unique primes and their frequency. One difference in the design is that LW had
a "noise" condition, in addition to the congruent/incongruent conditions.

```{r}
sort(table(lw1$prime, useNA = "ifany"))
```


Experimental cells
------------------

What is the proportion of items in each experimental cell

```{r}
lw1 %>%
  mutate(totalN = nrow(lw1)) %>%
  group_by(pic_presence, cue, condition, totalN) %>%
  summarise(N = n()) %>%
  mutate(prop = round(N / totalN, 3)) %>%
  arrange(pic_presence, cue, condition) %>%
  kable
```

By the very nature of the task, the two experimental factors ObjectPresence and
CueType cannot be fully crossed, because there is no "congruent" cue when there
is no object.

```{r}
with(lw1, addmargins(table(condition, pic_presence)))
```



Individual items
----------------

Now let's look at all prime-target combinations -- there is more imbalance than
in Markus's case:

```{r}
lw1 %>%
  group_by(prime, target, pic_presence, cue, condition) %>%
  count() %>%
  arrange(prime, target, pic_presence, cue, condition) %>%
  kable
```
